Limited evidence for widespread heterogeneity in psychology

### 2018-07-08 writing/appendixA-heterogeneity.rmd
Added results for simulation sensitivity check of effect size

#Summary of results. These are incorporated into the main table in the paper, see tables.rmd
zero <- dens %>% 
  bind_rows(.id = "effect") %>% 
  group_by(effect, tau_index) %>% 
  summarize(power = mean(Qp <= 0.05)) %>% #Estimate power/type 1 error for each tau level and effect
  ungroup() %>% 
  tidyr::spread(key = tau_index, value = power) %>% #prep for table
  rename(zero = '1', small = '2', medium = '3', large = '4') %>% 
  select(-effect)

@@

#Summary of results.
medium <- dens2 %>% 
  bind_rows(.id = "effect") %>% 
  group_by(effect, tau_index) %>% 
  summarize(power = mean(Qp <= 0.05)) %>% #Estimate power/type 1 error for each tau level and effect
  ungroup() %>% 
  tidyr::spread(key = tau_index, value = power) %>% #prep for table
  rename(zero = '1', small = '2', medium = '3', large = '4')

@@

Estimated type I error and power for zero/small/medium/large heterogeneity when simulated with a true effect size of zero vs. medium are shown for each effect in Table A1. In both cases type I error is consistently conservative, as compared to the expected 5% error rate. With a an effect size of zero, average power was `r 100*round(mean(zero$small), digits = 2)`/`r 100*round(mean(zero$medium), digits = 2)`/`r 100*round(mean(zero$large), digits = 2)`% for small/medium/large heterogeneity respectively. The corresponding power with an average effect size was `r 100*round(mean(zero$small), digits = 2)`, `r 100*round(mean(zero$medium), digits = 2)`, and `r 100*round(mean(zero$large), digits = 2)`%. To conclude, changing true effect size in our simulations from zero to medium has little impact on estimated type I error and power.

@@

_Type I error and power for data with a true effect size of zero vs. 'medium'_

@@

**Note.** Column headers indicate the degree of simulated heterogeneity where Zero = simulated type 1 error, and the other headers represent simulated power under small/medium/large heterogeneity ($I^2$ = 25/50/75%) respectively. Code to reproduce table: osf.io/bsjhu/

>>>>

#Summary of results. These are incorporated into the main table in the paper, see tables.rmd
zero <- dens %>% 
  bind_rows(.id = "effect") %>% 
  group_by(effect, tau_index) %>% 
  summarize(power = mean(Qp <= 0.05)) %>% #Estimate power/type 1 error for each tau level and effect
  ungroup() %>% 
  tidyr::spread(key = tau_index, value = power) %>% #prep for table
  rename(Zero = '1', Small = '2', Medium = '3', Large = '4', Effect = effect)

@@

#Summary of results.
medium <- dens2 %>% 
  bind_rows(.id = "effect") %>% 
  group_by(effect, tau_index) %>% 
  summarize(power = mean(Qp <= 0.05)) %>% #Estimate power/type 1 error for each tau level and effect
  ungroup() %>% 
  tidyr::spread(key = tau_index, value = power) %>% #prep for table
  rename(Zero = '1', Small = '2', Medium = '3', Large = '4') %>% 
  select(-effect)

@@

Estimated type I error and power for zero/small/medium/large heterogeneity when simulated with a true effect size of zero vs. medium are shown for each effect in Table A1. In both cases type I error is approximately nominal, as compared to the expected 5% error rate. With a an effect size of zero, average power was `r 100*round(mean(zero$Small), digits = 2)`%/`r 100*round(mean(zero$Medium), digits = 2)`%/`r 100*round(mean(zero$Large), digits = 2)`% for small/medium/large heterogeneity respectively. The corresponding power with an average effect size was `r 100*round(mean(zero$Small), digits = 2)`%/ `r 100*round(mean(zero$Medium), digits = 2)`%/`r 100*round(mean(zero$Large), digits = 2)`%. To conclude, changing true effect size in our simulations from zero to medium had little to no impact on estimated type I error and power.

@@

_Type I error and power to detect four heterogeniety levels under two true effect sizes_

@@

**Note.** Column headers indicate the degree of simulated heterogeneity where Zero = simulated type 1 error, and the other headers represent simulated power under small/medium/large heterogeneity ($I^2$ = 25/50/75%) respectively. Medium effect size is as defined by Cohen (1988). Code to reproduce table: osf.io/bsjhu/

### 2018-07-08 writing/draft-heterogeneity.rmd
Added results for simulation sensitivity check of effect size

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no

@@

As our concern was heterogeneity, for convenience we set the true effect size to zero in our simulations. This should not affect the results for correlations or mean differences, as estimates of effect size and heterogeneity for these measures are statistically independent. For standardized mean differences we expect negligible effects on the results, because while these estimates of effect size are positively correlated to their standard errors, the within study variance $\sigma^2$ was kept constant across studies. Nonetheless, as a sensitivity test we also ran the analysis with 'medium' effect sizes [@Cohen1988], see Appendix A (osf.io/bsjhu/).

@@

Estimated type I error and power for zero/small/medium/large heterogeneity are shown for each effect in Table 3. In all cases the type I error is conservative, as compared to the expected 5% error rate. Power to detect small heterogeneity was low, ranging between `r 100*round(min(het$small), digits = 2)`% and `r 100*round(max(het$small), digits = 2)`%, with an average of `r 100*round(mean(het$small), digits = 2)`%. Power to detect medium heterogeneity was generally very good, with an average of `r 100*round(mean(het$medium), digits = 2)`% power, but goes down to as low as `r 100*round(min(het$medium), digits = 2)`% for several effects with low K. Power to detect strong heterogeneity was excellent across the board. To conclude, even though for most projects the number of included studies (median `r quantile(sort(het$k), .5)[[1]]`) and number of participants (median 102 per study) was relatively large, only power to detect medium or larger heterogeneity was good to excellent, whereas power to detect small heterogeneity was unacceptably low. Hence, even large multi-lab projects struggle to distinguish zero from small heterogeneity.

>>>>

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes

@@

As our concern was heterogeneity, for convenience we set the true effect size to zero in our simulations. This should not affect the results for correlations or mean differences, as estimates of effect size and heterogeneity for these measures are statistically independent. For standardized mean differences we expect negligible effects on the results, because while these estimates of effect size are positively correlated to their standard errors, the within study variance $\sigma^2$ was kept constant across studies. As a sensitivity test we ran the analysis with 'medium' effect sizes [@Cohen1988] and found the same average power at the different heterogeneity levels, see Appendix A (osf.io/bsjhu/).

@@

Estimated type I error and power for zero/small/medium/large heterogeneity are shown for each effect in Table 3. In all cases the type I error is approximately nominal, as compared to the expected 5% error rate. Power to detect small heterogeneity was low, ranging between `r 100*round(min(het$small), digits = 2)`% and `r 100*round(max(het$small), digits = 2)`%, with an average of `r 100*round(mean(het$small), digits = 2)`%. Power to detect medium heterogeneity was generally very good, with an average of `r 100*round(mean(het$medium), digits = 2)`% power, but goes down to as low as `r 100*round(min(het$medium), digits = 2)`% for several effects with low K. Power to detect strong heterogeneity was excellent across the board. To conclude, even though for most projects the number of included studies (median `r quantile(sort(het$k), .5)[[1]]`) and number of participants (median 102 per study) was relatively large, only power to detect medium or larger heterogeneity was good to excellent, whereas power to detect small heterogeneity was unacceptably low. Hence, even large multi-lab projects struggle to distinguish zero from small heterogeneity.
